---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = TRUE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "150%"
)
```

# Under-Five Child Mortality Estimation using the R Package `u5mr`

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

`u5mr` is a open-source R package for estimating the child mortality. Current implementation includes the Trussell version of the Brass method using the Coale-Demeny model life tables and supporting datasets of coefficients and automatic interpolating values between probabilities of dying at a certain age and model tables.

## Installation

To download the developmental version of the u5mr package, use the code below. 

``` r
# install.packages("devtools")
devtools::install_github("myominnoo/u5mr")
```

## Usage

The first example is using Bangladesh survey data and model South of the Coale-Demeny life table. 

```{r demo1}
library(u5mr)

## Using Bangladesh survey data to estimate child mortality
data("bangladesh")
bang_both <- u5mr_trussell(bangladesh, sex = "both", model = "south", svy_year = 1974.3)
bang_both
bang_male <- u5mr_trussell(bangladesh, child_born = "male_born",
                 child_dead = "male_dead", sex = "male",
                 model = "south", svy_year = 1974.3)
bang_male
bang_female <- u5mr_trussell(bangladesh, child_born = "female_born",
                 child_dead = "female_dead", sex = "female",
                 model = "south", svy_year = 1974.3)
bang_female

## plotting all data points
with(bang_both,
    plot(year, q5, type = "b", pch = 19,
         ylim = c(0, .45),
         col = "black", xlab = "Reference date", ylab = "u5MR",
         main = paste0("Under-Five mortality, q(5) in Bangladesh, estimated\n",
                       "using model South and the Trussell version of the Brass method")))
with(bang_both, text(year, q5, agegrp, cex=0.65, pos=3,col="purple"))
with(bang_male, lines(year, q5, pch = 18, col = "red", type = "b", lty = 2))
with(bang_female,
    lines(year, q5, pch = 18, col = "blue", type = "b", lty = 3))
legend("bottomright", legend=c("Both sexes", "Male", "Female"),
      col = c("Black", "red", "blue"), lty = 1:3, cex=0.8)

```

Below, the second example is demonstrated using Panama survey data and model West.  

```{r demo2}
## Using panama survey data to estimate child mortality
data("panama")
pnm_both <- u5mr_trussell(panama, sex = "both", model = "west", svy_year = 1976.5)
pnm_both
pnm_male <- u5mr_trussell(panama, child_born = "male_born",
                child_dead = "male_dead", sex = "male",
                model = "west", svy_year = 1976.5)
pnm_male
pnm_female <- u5mr_trussell(panama, child_born = "female_born",
                child_dead = "female_dead", sex = "female",
                model = "west", svy_year = 1976.5)
pnm_female

## plotting all data points
with(pnm_both,
    plot(year, q5, type = "b", pch = 19,
        ylim = c(0, .2), col = "black", xlab = "Reference date", ylab = "u5MR",
         main = paste0("Under-Five mortality, q(5) in Panama, estimated\n",
                       "using model West and the Trussell version of the Brass method")))
with(pnm_both, text(year, q5, agegrp, cex=0.65, pos=3,col="purple"))
with(pnm_male,
    lines(year, q5, pch = 18, col = "red", type = "b", lty = 2))
with(pnm_female,
    lines(year, q5, pch = 18, col = "blue", type = "b", lty = 3))
legend("bottomleft", legend=c("Both sexes", "Male", "Female"),
      col = c("Black", "red", "blue"), lty = 1:3, cex=0.8)
```

## Bug Reports / Change Requests

If you encounter a clear bug, please file an issue with a minimal
reproducible example on
[GitHub](https://github.com/myominnoo/u5mr/issues). 

## Getting Help 
For questions and other discussion, please directly email me [dr.myominnoo@gmail.com](mailto::dr.myominnoo@gmail.com).

## Citation

To cite the u5mr package in publications use
```
  @Manual{Myo2020space,
    title = {Under-Five Child Mortality Estimation using the R Package u5mr},
    author = {Myo Minn Oo},
    year = {2021}
  }
```

-----

Please note that this project is looking for contributors. By
participating in this project, you agree to abide by its terms with
[Contributor Code of
Conduct](https://www.contributor-covenant.org/version/1/0/0/code-of-conduct/),
version 1.0.0, available at
<https://www.contributor-covenant.org/version/1/0/0/code-of-conduct/>.
